---
title: "Identifying nests and swallows based on spatiotemporal wild boar data."
subtitle: "Project Works Patterns and Trends in Environmental Data"
author: "Yves Gubelmann and Tobias Wildhaber"
date: "4 July, 2021"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: darkly
    highlight: tango
    code_folding: hide
bibliography: references.bib
---

<!-- You can add  your R Code with Code chunks-->

```{r chunk settings, echo = TRUE, warning=FALSE, message=FALSE}

# You can set chunk options individually per code chunk, as was done with this
# code chunk.

# echo = FALSE           hides the code from the generated output
# warning = FALSE        hides warnings from the generated output
# message = FALSE        hides messages from the generated output

# To set the setting for all code chunks, use the following code:

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```

# Abstract

<!-- the following is just a placeholder text, remove it!-->

Tobias

# Research Questions

(yves) Our research is based on data from the ZHAW research project "Prevention of Wild Boar Damage in Agriculture"[@suter2018]. With this dataset as a basis, we want to answer the following research questions:

-   How can resting sites be modelled based on spatiotemporal movement data?

-   Can locations frequently visited by wild boar be determined efficiently and effectively in the provided research data?

-   Can these regular resting places be clearly defined as nests or wallows?

# Material and Methods

## Data

The provided data contains spatio temporal datasets of 39 wildboars which were equipped with the GPS collar VERTEX PLUS [@vectronicaerospace2021]. Some wildboars were equipped multiple time which results in 46 unique datasets. The data was taken between 2014-05-28 and 2016-10-18 on a irregular basis. This results in

We will mainly use data that we have already used in the exercises. The basis is the wild boar movement
dataset and its metadata. In addition, we will also use the raster and vector data provided on the environment (field and vegetation type). Additional data on land coverage may be required (e.g. area statistics, vegetation height), which would be obtained from swisstopo.

Yves

Datenexploration, Rohdaten beschreiben, Plots aus Exploration

Feldaufnahmen berücksichtigen

## Analytical Concepts

Yves

Stepmean, Segmentation,

Anhand von Literatur Kriterien festgelegt für Nester und Suhlen

Sampling Regime (nur 15 Minuten Intervalle)

## R Concecpts

Tobias

Umsetzungskonzepte in R

Packages mit deren Funktion

# Results / Products

Tobias

Boxplot mit Stepmean, Exemplarische Plots nach Tieren, Raster graphs, Anteile Site Types & Moving / Resting,

Text

# Discussion

Yves

Warum so viele none und auch both? Evtl. Tabelle der Anteile hier bringen.

## Risk analysis / Problems

Keine Validierung möglich, Literaturangaben zur Differenzierung nach N/W unzureichend. Intervalle der Messungen diskutieren, nur 15 Min berücksichtigt, dadurch Datenvelrust

<!-- the following is just a placeholder text, remove it!-->

Text

# References

Tobias

Zotero und so

# Annex

R Script

```{r source code, echo=TRUE}

# source("scripts/analysis.r")

```

[@baettig2021]
