---
title: "Identifying nests and swallows based on spatiotemporal wild boar data."
subtitle: "Project Works Patterns and Trends in Environmental Data"
author: "Yves Gubelmann and Tobias Wildhaber"
date: "4 July, 2021"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: darkly
    highlight: tango
    code_folding: hide
bibliography: references.bib
---

<!-- You can add  your R Code with Code chunks-->

```{r chunk settings, echo = TRUE, warning=FALSE, message=FALSE}

# You can set chunk options individually per code chunk, as was done with this
# code chunk.

# echo = FALSE           hides the code from the generated output
# warning = FALSE        hides warnings from the generated output
# message = FALSE        hides messages from the generated output

# To set the setting for all code chunks, use the following code:

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```

# Abstract

<!-- the following is just a placeholder text, remove it!-->

Tobias

# Research Questions

(yves) Our research is based on data from the ZHAW research project "Prevention of Wild Boar Damage in Agriculture"[@suter2018]. With this dataset as a basis, we want to answer the following research questions:

-   How can resting sites be modelled based on spatiotemporal movement data?

-   Can locations frequently visited by wild boar be determined efficiently and effectively in the provided research data?

-   Can these regular resting places be clearly defined as nests or wallows?

# Material and Methods

## Data

The provided data contains spatio temporal datasets of 39 wildboars which were equipped with the GPS collar VERTEX PLUS [@vectronicaerospace2021]. Some wildboars were equipped multiple times which results in 46 unique datasets. The data was taken between 2014-05-28 and 2016-10-18 around the conservation zone Fanel which belongs to the cantons Bern and Neuenburg. The sampling regime varies widely in duration and interval. The vast majority of the recorded wild boar data has a timelag of 900 seconds between to datapoints (see *Figure 1*).

-\> Sampling Regime bild

Furthermore we added spatial information about the vegetation to the spatiotemporal wild boar data. This data was provided by a survey which classified the investigated area in forests, wetlands, settlement and agriculture. If a lot of land was used for agriculture, information about the planted crop is also given. In total the survey evaluated 5154 lots of land, covering an area of 120 km\^2. The evaluated area consists of 63 % agriculture, 22 % forest, 9 % settlement and 6 % wetlands. [@suter2010]

Yves

Datenexploration, Rohdaten beschreiben, Plots aus Exploration

Feldaufnahmen berücksichtigen

## Analytical Concepts

In a prelimary step we enriched the given data by some important movement parameters. For every data point we calculated steplength and timelag to the next point.

Wild boars share a common movement pattern when spending time at wallows and nests: they stick to a spot for a certain while. Distinguishing this behavior from other (feeding, fleeing, strolling) builds the basement of our analytical concept. For this we calculated the mean distance within a moving window of six data points (stepmean). This variable allowed us to apply segmentation on our wild boar data and define stops and moves. In our work we called it *movement status* (“resting” or “moving”).

In a second phase we defined the threshold of stepmean. For this stepmean of three wild boars were plotted in a histogram. We expected to see a clear decline in the first bins of stepmean since wild boars gather at spots and move often in a slow manner.

Setting the threshold seems reasonable at the point where stepmean flattens in a distinct manner (see histograms). In our samples this threshold can be anywhere between 30 to 50 meter. 


This was followed by finding a meaningful parameter that allows us to allocate 


### Finding Stepmean

be based on predefined movement properties (e.g. speed, step length) which need to be readjusted multiple times and reviewed in a critical manner. The resulting trajectories will then be annotated with a status  and joined with land use type to provide further context. Finally, for each resting trajectory a convex hull will be calculated to visualize the results.

Stepmean, Segmentation,

Anhand von Literatur Kriterien festgelegt für Nester und Suhlen

Sampling Regime (nur 15 Minuten Intervalle)

## R Concecpts

Tobias

Umsetzungskonzepte in R

Packages mit deren Funktion

# Results / Products

Tobias

Boxplot mit Stepmean, Exemplarische Plots nach Tieren, Raster graphs, Anteile Site Types & Moving / Resting,

Text

# Discussion

Yves

Warum so viele none und auch both? Evtl. Tabelle der Anteile hier bringen.

## Risk analysis / Problems

Keine Validierung möglich, Literaturangaben zur Differenzierung nach N/W unzureichend. Intervalle der Messungen diskutieren, nur 15 Min berücksichtigt, dadurch Datenvelrust

<!-- the following is just a placeholder text, remove it!-->

Text

# References

Tobias

Zotero und so

# Annex

R Script

```{r source code, echo=TRUE}

# source("scripts/analysis.r")

```

[@baettig2021]
