---
title: "Identifying nests and swallows based on spatiotemporal wild boar data."
subtitle: "Project Works Patterns and Trends in Environmental Data"
author: "Yves Gubelmann and Tobias Wildhaber"
date: "4 July, 2021"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: darkly
    highlight: tango
    code_folding: hide
bibliography: references.bib
---

<!-- You can add  your R Code with Code chunks-->

```{r chunk settings, echo = TRUE, warning=FALSE, message=FALSE}

# You can set chunk options individually per code chunk, as was done with this
# code chunk.

# echo = FALSE           hides the code from the generated output
# warning = FALSE        hides warnings from the generated output
# message = FALSE        hides messages from the generated output

# To set the setting for all code chunks, use the following code:

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```

# Abstract

<!-- the following is just a placeholder text, remove it!-->

Tobias

# Research Questions

(yves) Our research is based on data from the ZHAW research project "Prevention of Wild Boar Damage in Agriculture"[@suter2018]. With this dataset as a basis, we want to answer the following research questions:

-   How can resting sites be modelled based on spatiotemporal movement data?

-   Can locations frequently visited by wild boar be determined efficiently and effectively in the provided research data?

-   Can these regular resting places be clearly defined as nests or wallows?

# Material and Methods

## Data

The provided data contains spatio temporal datasets of 39 wildboars which were equipped with the GPS collar VERTEX PLUS [@vectronicaerospace2021]. Some wildboars were equipped multiple times which results in 46 unique datasets. The data was taken between 2014-05-28 and 2016-10-18 around the conservation zone Fanel which belongs to the cantons Bern and Neuenburg. The sampling regime varies widely in duration and interval. 

-\> Sampling Regime bild

Furthermore we added spatial information about the vegetation to the spatiotemporal wild boar data. This data was provided by a survey which classified the investigated area in forests, wetlands and agriculture. If a lot of land was used for agriculture, information about the planted crop is also provided.

Yves

Datenexploration, Rohdaten beschreiben, Plots aus Exploration

Feldaufnahmen berücksichtigen

## Analytical Concepts



Stepmean, Segmentation,

Anhand von Literatur Kriterien festgelegt für Nester und Suhlen

Sampling Regime (nur 15 Minuten Intervalle)

## R Concecpts

Tobias

Umsetzungskonzepte in R

Packages mit deren Funktion

# Results / Products

Tobias

Boxplot mit Stepmean, Exemplarische Plots nach Tieren, Raster graphs, Anteile Site Types & Moving / Resting,

Text

# Discussion

Yves

Warum so viele none und auch both? Evtl. Tabelle der Anteile hier bringen.

## Risk analysis / Problems

Keine Validierung möglich, Literaturangaben zur Differenzierung nach N/W unzureichend. Intervalle der Messungen diskutieren, nur 15 Min berücksichtigt, dadurch Datenvelrust

<!-- the following is just a placeholder text, remove it!-->

Text

# References

Tobias

Zotero und so

# Annex

R Script

```{r source code, echo=TRUE}

# source("scripts/analysis.r")

```

[@baettig2021]
